Confirmation of Wide-Field Signatures in Redshifted 21~cm Power Spectra

### 2015-06-02 ms.tex
Modify Abstract
From bc89703:
============

In ms.tex,

(*) Modify Abstract

We confirm the recently predicted ``pitchfork'' foreground signature in power spectra of high-redshift 21~cm measurements, wherein the interferometer is sensitive to large-scale structure on all baselines. This is due to the inherent response of a wide-field instrument and is characterized by enhanced power from foreground emission in Fourier modes adjacent to those considered to be most sensitive to the cosmological H~{\sc i} signal. In our recent paper, many signatures in the simulation which predicted this feature were validated against Murchison Widefield Array data but this key element was at or below the noise level. In this paper, we improve the data sensitivity through coherent averaging of 12 independent snapshots with identical instrument settings, and provide the first confirmation of the prediction with a signal-noise ratio $>10$. This wide-field effect can by mitigated by careful antenna design which suppresses sensitivity towards the horizon. With future instruments such as the Hydrogen Epoch of Reionization Array and the Square Kilometre Array, based on simple scaling relations, we estimate a suppression of $\gtrsim 40$~dB is required near the horizon for signal detection in cosmological H~{\sc i} modes in analyses that employ {\it foreground avoidance} techniques.

@@

We approached this issue in Paper~I via simulations of different antenna apertures, namely, a dipole, a phased array, and a dish. Among these, a dish provides the best quality as evaluated from a foreground contamination viewpoint. With new instruments such as HERA and SKA on the horizon, we extend this discussion to further emphasize the need to comprehensively study the effect of their antenna apertures on foreground contamination.

@@

With the SKA at low frequencies, deploying ``stations'' of $\sim$~35~m diameter that act as aperture arrays consisting of 256 pseudo-randomly placed vertical log-periodic dipole antennas is being considered\footnote{Documents at \url{https://www.skatelescope.org/key-documents/} on SKA re-baselining}$^,$\footnote{SKA memo ``Station Response and Imaging Performance of LFAA: 100 MHz to 600 MHz'' by Razavi-Ghods et al.}. The voltage beam of the station will be obtained by a phased addition of the dipole responses. Since the power pattern of a baseline consisting of two stations will be effectively a product of two voltage beams, it will tend to homogenize the fluctuations in the power patterns of individual stations. Even with unconstrained, and hence sub-optimal, realizations of these phased array stations, we estimate the typical horizon response of these beams to be $\gtrsim$~30--40~dB lower than at zenith, also $\sim 20$~dB lower than the MWA beam. Constrained optimization studies of antenna layouts in the stations are under active study and could yield even better responses.

>>>>

We confirm our recent prediction of the ``pitchfork'' foreground signature in power spectra of high-redshift 21~cm measurements, wherein the interferometer is sensitive to large-scale structure on all baselines. This is due to the inherent response of a wide-field instrument and is characterized by enhanced power from foreground emission in Fourier modes adjacent to those considered to be most sensitive to the cosmological H~{\sc i} signal. In our recent paper, many signatures from the simulation which predicted this feature were validated against Murchison Widefield Array (MWA) data but this key element was at or below the noise level. In this paper, we improve the data sensitivity through coherent averaging of 12 independent snapshots with identical instrument settings, and provide the first confirmation of the prediction with a signal-noise ratio $>10$. This wide-field effect can by mitigated by careful antenna design which suppresses sensitivity towards the horizon. Simple models for antenna apertures proposed for future instruments such as the Hydrogen Epoch of Reionization Array and the Square Kilometre Array indicate they should suppress foreground contamination by $\sim 40$~dB relative to the MWA. However, detailed models are necessary to ensure the contamination stays below the expected cosmological signal.

@@

We approached this issue in Paper~I via simulations of different antenna apertures, namely, a dipole, a $4\times 4$ phased array, and a dish. Among these, a dish provides the best quality as evaluated from a foreground contamination viewpoint. With new instruments such as HERA and SKA on the horizon, we extend this discussion to further emphasize the need to comprehensively study the effect of their antenna apertures on foreground contamination.

@@

With the SKA at low frequencies, deploying ``stations'' of $\sim$~35~m diameter that act as aperture arrays consisting of 256 pseudo-randomly placed vertical log-periodic dipole antennas is being considered\footnote{Documents at \url{https://www.skatelescope.org/key-documents/} on SKA re-baselining}$^,$\footnote{SKA memo ``Station Response and Imaging Performance of LFAA: 100 MHz to 600 MHz'' by Razavi-Ghods et al.}. The voltage beam of the station will be obtained by a phased addition of the dipole responses. Since the power pattern of a baseline consisting of two stations will be effectively a product of two voltage beams, it will tend to homogenize the fluctuations in the power patterns of individual stations. Even with unconstrained, and hence sub-optimal, realizations of these phased array stations, we estimate the typical horizon response of these beams to be $\gtrsim$~30--40~dB lower than at zenith, also $\sim 20$~dB lower than the MWA beam. Constrained optimization studies of antenna layouts of stations are under active study and could yield even better responses.
